The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-04-23 Paper/jaccardbaetica.tex
first correction xavi

\documentclass[review]{elsarticle}
\usepackage{hyperref,lineno}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\modulolinenumbers[5]

@@

%This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
This paper aims to understand the production dynamics in connection with different areas within the Roman Empire. Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres. To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}. Stamps will be used to identify economic patterns by analysing their similarity. If workshops and provinces share stamps with similar traits, then we can identify connections. By contrast, if we do not detect similar stamps between workshops, then workshops worked independently.

@@

\subsection{The Amphoric production in Baetica province}

@@

\subsection{A potential indicator of Roman economy: Dressel 20 stamps}

@@

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}). In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be affected by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.

Therefore, this study proposes an evolutionary baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.

@@

@@

\subsubsection{Material}

We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province. The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}). However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information. Consequently, we discard integrate the fragmented stamps in our dataset. We finally filter a total sample of 987 stamps composed of 130 different stamps from 81 workshops.

@@

\subsubsection{Exploratory Analysis}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
\caption{Distribution of the number of stamps}
\label{stamps}
\end{figure}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequency}
\caption{Distribution of the number of different code stamps for each area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue).}
\label{frequency}
\end{figure}

@@

\subsubsection{Britannia}

\subsubsection{Material}

@@

\subsubsection{Germania}

The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto. The advance of the Roman legions toward \textit{Gallia} opened a path through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
remesal_baetica_2002}.

It is thought that this route would have been initiated previously through terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of \textit{Germania} and those of \textit{Britannia} \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.

However, the Roman studies about the presence of Dressel 20 amphorae in textit{Germania} has not had the same repercussion as the rest of the province. This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.

\subsubsection{Material}

@@

%hablar sobre Horn (Morisita)hablar sobre que el sample no estaba uniforme por eso usamos el morisita horn

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/dendro}
\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in \textit{Baetica} area. Colours are represented by areas divided into \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
\label{dendro}
\end{figure}

@@

\subsubsection{Britannia}

The results show a correlation similar to the obtained in \textit{Baetica} province. However, the similarity is less pronounced as we observed in the dendrogram in Fig. \ref{britmap}

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/dendrobrit5.pdf}
\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Britannia}. Colours are represented by type of sites divided into military sites (red), civil sites (green) and no specific (grey)}
\label{britmap}
\end{figure}

@@

\subsubsection{Germania}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/dendroger5.pdf}
\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Germania}. Colours are represented by types of sites divided into military sites (red), civil sites (green) and no specific (grey)}
\label{germap}
\end{figure}

>>>>

\documentclass[review]{elsarticle}
\usepackage{hyperref,lineno}
\usepackage{wrapfig}
\usepackage{lscape}
\usepackage{rotating}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\modulolinenumbers[5]

@@

%This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
\xavi{este aim es demasiado general}
This paper aims to understand the production dynamics in connection with different areas within the Roman Empire. Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres. To do this, an population approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}. Stamps will be used to identify economic patterns by analysing their similarity. If workshops and provinces share stamps with similar traits, then we can identify connections. By contrast, if we do not detect similar stamps between workshops, then workshops worked independently.

%\xavi{esta frase tendría que ir al final de la Intro; si pones subsecciones entonces queda muy raro creo}

@@

\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}

\section{The Amphoric production in Baetica province}

@@

\section{Britannia and Germania}

%\xavi{Esta parte más histórica está chachi pero creo que debería estar en la Intro. Ahí podrías explicar la organización, y a partir de estas fuentes escritas plantear preguntas e hipótesis que justifiquen el paper y tu análisis usando arqueología.}

The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto. The advance of the Roman legions toward \textit{Gallia} opened a path through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
remesal_baetica_2002}. It is thought that this route would have been initiated previously through terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of \textit{Germania} and those of \textit{Britannia} \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.

However, the Roman studies about the presence of Dressel 20 amphorae in \textit{Germania} has not had the same repercussion as the rest of the province. This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.

\section{A potential indicator of Roman economy: Dressel 20 stamps}

@@

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}). In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province. 
%\xavi{affected? No sería "identified"? Es decir, la correlación es la evidencia, pero la causa sería la interacción económica...no?}

Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance. 
%\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}

@@

%\xavi{Yo aplicaría aquí el feedback de los mapas para lo del libro...no creo que colorearlo basándose en el DEM aporte demasiado no?}

@@

\section{Method and Analysis}

We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province. The stamp database was compiled by CEIPAC group \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}). However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information. Consequently, we discard integrate the fragmented stamps in our dataset. We finally filter a total sample of 987 stamps composed of 130 different stamps from 81 workshops.

@@

%\xavi{a ver, esto queda raro...si esto es una exploración de la base de datos entonces tiene que ir dentro de "Materials", así mientras vas comentando los sesgos y cosas generales de los datos (como soporte, pero no hace falta subsubsección "EDA")}

@@

\xavi{quieres decir que esto es una power base? yo creo que no, no es tan bestia no? Yo lo matizaría}

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
\caption{Histogram with the frequency of the number of stamps per workshop. The distribution is widely diverse with most stamps concentrated in one workshop}
\label{stamps}
\end{figure}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequency}
\caption{Distribution of the number of different code stamps for each area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
\label{frequency}
\end{figure}

@@

@@

@@

\subsection{Hieralchical clustering}

\maria{que pongo aquí}

%\xavi{pero este es el único método que usas?}

@@

\begin{sidewaysfigure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/dendro}
\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in \textit{Baetica} area. Colours are represented by areas divided into \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
\label{dendro}
\end{sidewaysfigure}

@@

The results show a similarity with the results obtained in \textit{Baetica} province. However, the similarity is less pronounced as we observed in the dendrogram in Fig. \ref{britmap}

%\xavi{no tengo claro que le puedas decir correlación, porque eso es un valor específico entre 2 distribuciones numéricas. Te refieres a similaridad?}

\begin{sidewaysfigure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/dendrobrit5.pdf}
\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Britannia}. Colours are represented by type of sites divided into military sites (red), civil sites (green) and no specific (grey)}
\label{britmap}
\end{sidewaysfigure}

@@

@@

\begin{sidewaysfigure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/dendroger5.pdf}
\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Germania}. Colours are represented by types of sites divided into military sites (red), civil sites (green) and no specific (grey)}
\label{germap}
\end{sidewaysfigure}
